---
title: "Assignment 3 - Part 1 - Voice In Schizophrenia"
author: "Riccardo Fusaroli"
date: "August 09, 2019"
output: html_document
---
I DAG - COLLECT CODE !!!!

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Assignment 3 - Part 1 - Assessing voice in schizophrenia

Individuals with schizophrenia (SCZ) tend to present voice atypicalities. Their tone is described as "inappropriate" voice, sometimes monotone, sometimes croaky. This is important for two reasons. First, voice could constitute a direct window into cognitive, emotional and social components of the disorder, thus providing a cheap and relatively non-invasive way to support the diagnostic and assessment process (via automated analyses). Second, voice atypicalities play an important role in the social impairment experienced by individuals with SCZ, and are thought to generate negative social judgments (of unengaged, slow, unpleasant interlocutors), which can cascade in more negative and less frequent social interactions.

Several studies show *significant* differences in acoustic features by diagnosis (see meta-analysis in the readings), but we want more. We want to know whether we can diagnose a participant only from knowing the features of their voice.

The corpus you are asked to analyse is a relatively large set of voice recordings from people with schizophrenia (just after first diagnosis) and matched controls (on gender, age, education). Each participant watched several videos of triangles moving across the screen and had to describe them (so you have several recordings per person). We have already extracted the pitch once every 10 milliseconds as well as several duration related features (e.g. number of pauses, etc).

N.B. For the fun of it, I threw in data from 3 different languages: 1) Danish (study 1-4); 2) Mandarin Chinese (Study 5-6); 3) Japanese (study 7). Feel free to only use the Danish data, if you think that Mandarin and Japanese add too much complexity to your analysis.

In this assignment (A3), you will have to discuss a few important questions (given the data you have). More details below.

*Part 1 - Can we find a difference in acoustic features in schizophrenia?*
1) Describe your sample 
number of studies, number of participants, age, gender, clinical and cognitive features of the two groups. Furthemore, critically assess whether the groups (schizophrenia and controls) are balanced. N.B. you need to take studies into account.


2) Describe the acoustic profile of a schizophrenic voice: which features are different?
E.g. People with schizophrenia tend to have high-pitched voice, and present bigger swings in their prosody than controls. N.B. look also at effect sizes. How do these findings relate to the meta-analytic findings?

3) Discuss the analysis necessary to replicate the meta-analytic findings
Look at the results reported in the paper (see meta-analysis in the readings) and see whether they are similar to those you get.
    3.1) Check whether significance and direction of the effects are similar
    3.2) Standardize your outcome, run the model and check whether the beta's is roughly matched (matched with hedge's g)
which fixed and random effects should be included, given your dataset? E.g. what about language and study, age and gender? Discuss also how studies and languages should play a role in your analyses. E.g. should you analyze each study individually? Or each language individually? Or all together? Each of these choices makes some assumptions about how similar you expect the studies/languages to be.
*Note* that there is no formal definition of replication (in statistical terms).


Your report should look like a methods paragraph followed by a result paragraph in a typical article (think the Communication and Cognition paper)

*Part 2 - Can we diagnose schizophrenia from voice only?*
1) Discuss whether you should you run the analysis on all studies and both languages at the same time
You might want to support your results either by your own findings or by that of others
2) Choose your best acoustic feature from part 1. How well can you diagnose schizophrenia just using it?
3) Identify the best combination of acoustic features to diagnose schizophrenia using logistic regression.
4) Discuss the "classification" process: which methods are you using? Which confounds should you be aware of? What are the strength and limitation of the analysis?

Bonus question: Logistic regression is only one of many classification algorithms. Try using others and compare performance. Some examples: Discriminant Function, Random Forest, Support Vector Machine, Penalized regression, etc. The packages caret and glmnet provide them. Tidymodels is a set of tidyverse style packages, which take some time to learn, but provides a great workflow for machine learning.

## Learning objectives
- Critically design, fit and report multilevel regression models in complex settings
- Critically appraise issues of replication


## Overview of part 1
In the course of this part 1 of Assignment 3 you have to:
- combine the different information from multiple files into one meaningful dataset you can use for your analysis. This involves: extracting descriptors of acoustic features from each pitch file (e.g. mean/median, standard deviation / interquartile range), and combine them with duration and demographic/clinical files
- describe and discuss your sample
- analyze the meaningful dataset to assess whether there are indeed differences in the schizophrenic voice and compare that to the meta-analysis

There are three pieces of data:

1- Demographic data (https://www.dropbox.com/s/e2jy5fyac18zld7/DemographicData.csv?dl=0). It contains

- Study: a study identifier (the recordings were collected during 6 different studies with 6 different clinical practitioners in 2 different languages)
- Language: Danish, Chinese and Japanese
- Participant: a subject ID
- Diagnosis: whether the participant has schizophrenia or is a control
- Gender
- Education
- Age
- SANS: total score of negative symptoms (including lack of motivation, affect, etc). Ref: Andreasen, N. C. (1989). The Scale for the Assessment of Negative Symptoms (SANS): conceptual and theoretical foundations. The British Journal of Psychiatry, 155(S7), 49-52.
- SAPS: total score of positive symptoms (including psychoses, such as delusions and hallucinations): http://www.bli.uzh.ch/BLI/PDF/saps.pdf
- VerbalIQ: https://en.wikipedia.org/wiki/Wechsler_Adult_Intelligence_Scale
- NonVerbalIQ: https://en.wikipedia.org/wiki/Wechsler_Adult_Intelligence_Scale
- TotalIQ: https://en.wikipedia.org/wiki/Wechsler_Adult_Intelligence_Scale

2. Articulation.txt (https://www.dropbox.com/s/vuyol7b575xdkjm/Articulation.txt?dl=0). It contains, per each file, measures of duration:
- soundname: the name of the recording file
- nsyll: number of syllables automatically inferred from the audio
- npause: number of pauses automatically inferred from the audio (absence of human voice longer than 200 milliseconds)
- dur (s): duration of the full recording
- phonationtime (s): duration of the recording where speech is present
- speechrate (nsyll/dur): average number of syllables per second
- articulation rate (nsyll / phonationtime): average number of syllables per spoken second
- ASD (speakingtime/nsyll): average syllable duration

3. One file per recording with the fundamental frequency of speech extracted every 10 milliseconds (excluding pauses): https://www.dropbox.com/sh/bfnzaf8xgxrv37u/AAD2k6SX4rJBHo7zzRML7cS9a?dl=0
- time: the time at which fundamental frequency was sampled
- f0: a measure of fundamental frequency, in Herz

NB. the filenames indicate:
- Study: the study, 1-6 (1-4 in Danish, 5-6 in Mandarin Chinese)
- D: the diagnosis, 0 is control, 1 is schizophrenia
- S: the subject ID (NB. some controls and schizophrenia are matched, so there is a 101 schizophrenic and a 101 control). Also note that study 5-6 have weird numbers and no matched participants, so feel free to add e.g. 1000 to the participant ID in those studies.
- T: the trial, that is, the recording ID for that participant, 1-10 (note that study 5-6 have more)

### Getting to the pitch data

You have oh so many pitch files. What you want is a neater dataset, with one row per recording, including a bunch of meaningful descriptors of pitch. For instance, we should include "standard" descriptors: mean, standard deviation, range. Additionally, we should also include less standard, but more robust ones: e.g. median, iqr, mean absoluted deviation, coefficient of variation. The latter ones are more robust to outliers and non-normal distributions.

Tip: Load one file (as a sample) and:
- write code to extract the descriptors
- write code to extract the relevant information from the file names (Participant, Diagnosis, Trial, Study)
Only then (when everything works) turn the code into a function and use map_df() to apply it to all the files.
See placeholder code here for help.

```{r}
pacman::p_load(tidyverse, stringr,purrr,lme4,ggpubr,effsize,MuMIn,caret)
```


```{r}
read_pitch <- function(filename) {
    # load data
    df <- read.delim(filename, sep = "\t", header = T ) #Because they are tab separated
    
    # parse filename to extract study, diagnosis, subject and trial
    Study = str_extract(filename,'Study.') %>% 
        str_remove('Study') 
    
    Diagnosis = str_extract(filename, 'D.') %>% 
        str_remove('D')
    
    Subject = str_extract(filename, 'S(\\d\\d\\d|\\d\\d)') %>%  #Look for S followed by 3 digits first, then S followed by 2 digits
        str_remove('S')
    
    # To find trial, we will specify the position of T (because some filenames include an end saying like "T12T13", and then we take the rest of the name)
    pos = str_locate(filename, 'T')[1] # Locating T, where we get a start and an end. We are only interested in the 'Start', the first value, so we index [1]
    split = length((str_split(filename, ''))[[1]]) # Split by 'nothing', so split everything that has nothing in between it = everything. Then, taking the length of that and indexing by 1 also. We take the length because that value will be the same as 'location of the end of the filename'
    
    Trial = substr(filename, pos, split) # Extracting from filename whatever is between pos (position of first occuring T) until the end of the filename ( substr(file, start, end).)
    
    Trial <- ifelse(str_detect(Trial, '_') == T, #if it has an underscore, remove _f0.txt, otherwise just .txt
                    str_remove(Trial, '_f0.txt'), 
                    str_remove(Trial, '.txt'))
    soundname <- ifelse(str_detect(filename, '_') == T,
                    str_remove(filename, '_f0.txt'), 
                    str_remove(filename, '.txt')) %>% str_remove('Pitch/')
    # extract pitch descriptors (mean, sd, iqr, etc)
    mean = mean(df$f0)
    sd = sd(df$f0)
    iqr = IQR(df$f0)
    median = median(df$f0)
    
    # combine all this data in one dataset
    data = c(soundname,filename, Study, Diagnosis, Subject, Trial, mean, sd, iqr, median)
    #data <- data %>% as.data.frame()
    #data <- data %>% t()
    #colnames(data) <- c("filename", "Study", "Diagnosis", "Subject", "Trial", "mean", "sd", "iqr", "median" )
    return(data)
}

# test it on just one file while writing the function
#test_data = read_pitch('Pitch/Study1D0S101T1_f0.txt')

# adding Pitch to the filenames (in front)
pastyp <- function(filename){
    filename <- paste('Pitch/', filename, sep = '')
    return(filename)
}
```

Reading in data:
```{r}
setwd('/Users/pernillebrams/Desktop/EM3/methods3_A3')

# my list
pitch_data = list.files(path = "Pitch/",pattern = ".txt") %>% 
    lapply(pastyp) %>% #adding Pitch to the filenames
    data.frame() %>% 
    purrr::map_df(read_pitch) %>% #applying read_pitch function to each element in the list
    t() %>% 
    data.frame()

# Fixing the column names
colnames(pitch_data) <- c('soundname',
                          'Filename',
                          'Study',
                          'Diagnosis', # 1 = skizo, 0 = control
                          'Participant',
                          'Trial',
                          'Mean',
                          'SD',
                          'IQR',
                          'Median')

# Fixing rownames
rownames(pitch_data) <- 1:nrow(pitch_data) #giving the rows numbers from 1 to total number of rows of pitch_dataframe

# Fixing some classes 
pitch_data$IQR <- as.numeric(pitch_data$IQR)
pitch_data$Mean <- as.numeric(pitch_data$Mean)
pitch_data$SD <- as.numeric(pitch_data$SD)
pitch_data$Median <- as.numeric(pitch_data$Median)
pitch_data$Participant <- as.numeric(pitch_data$Participant)

pitch_data$Study <- as.numeric(pitch_data$Study)
pitch_data$Diagnosis <- as.factor(pitch_data$Diagnosis)

#write_csv(pitch_data, "pitch_data_danish_only.csv")
```

# Loading Demographic
```{r}
# Loading more data, selecting only the Danish study
Dem_data <- read.csv("DemographicData.csv", sep = ';', header = T) %>% 
    filter(Language == "Danish")
    
# Making some stuff
Dem_data$Study <- as.numeric(Dem_data$Study)
Dem_data$Gender <- as.factor(Dem_data$Gender)
Dem_data$Diagnosis <- as.factor(Dem_data$Diagnosis)

        #so Control is
which(Dem_data$Diagnosis == 'Control') #from 1-130
which(Dem_data$Diagnosis == 'Schizophrenia') #from 131-264

####### GATHER THOSE TWO IN THE PIPES ####
```

# Loading Articulation data
```{r}

Arti_data <- read.delim('Articulation.txt', sep = ',', header = T)

```

### Now you need to merge demographic/clinical, duration and pitch data

```{r}
# Before merge, we want to add 1000 for Schizoprenia: (have 1-4 in pitch dem)
Dem_data$Participant <- ifelse(Dem_data$Diagnosis == "Schizophrenia", as.numeric(Dem_data$Participant)+1000, Dem_data$Participant)

pitch_data$Participant <- ifelse(pitch_data$Diagnosis == "1", as.numeric(pitch_data$Participant)+1000, pitch_data$Participant)
```

```{r}
# Now we merge them
merged_df <- merge(pitch_data, Arti_data, by = 'soundname')
merged_df <- merge(merged_df,Dem_data, by = 'Participant')

# The maximum mean is with subject 1040, of #780.1012 Hz
which.max(merged_df$Mean)
merged_df$Mean[1040] 

# Removing so we only have Danish studies
merged_df <- merged_df %>% filter(Study.x <5)

merged_df <- merged_df %>% filter(Study.y <5)

# The maximum mean is with subject 858, of 305.669 HZ
which.max(merged_df$Mean)
merged_df$Mean[858] 

# Select relevant variables
variable.names(merged_df)

df <- merged_df %>% select(
    c(
        Diagnosis.x,
        Diagnosis.y,
        Participant,
        SANS,
        SAPS,
        Gender,
        Study.y,
        Trial,
        npause,
        dur..s.,
        phonationtime..s.,
        IQR,
        speechrate..nsyll.dur.,
        Age,
        Mean,
        Median
    )
) %>%
    na.omit()

```

## Now we need to describe our sample

First look at the missing data: we should exclude all recordings for which we do not have complete data.
Then count the participants and recordinsgs by diagnosis, report their gender, age and symptom severity (SANS, SAPS and Social)
Finally, do the same by diagnosis and study, to assess systematic differences in studies.
I like to use group_by() %>% summarize() for quick summaries

```{r}

# # count participants
# # because some participants occur various times in the dataset, a subset of Unique ID's will be made here to find the actual age
# sub_sounddata <- sounddata %>%
#     group_by(ID) %>% 
#     slice(1)
# 
# sub_sounddata %>% group_by(Diagnosis.y) %>% summarise(mean(Age)) #Control: 23.1, Schizophrenia 23.0
# 
# sounddata %>% group_by(Study.x) %>% summarise(count=n()) #Study 1 = 683, Study 2 = 363
# sounddata %>% group_by(Gender) %>% summarise(mean(Age)) #F = 21.9, M = 23.7

```

In order to avoid collinearity, we constructed a heatmap of the correlations between the variables.
```{r}
# Defining the functions needed for the heatmap. We use correlation between variables as distance
reorder_cormat <- function(cormat) {
  dd <- as.dist((1 - cormat) / 2)
  hc <- hclust(dd)
  cormat <- cormat[hc$order, hc$order]
  }

# Get upper triangle of the correlation matrix
get_upper_tri <- function(cormat) {
  cormat[lower.tri(cormat)] <- NA
  return(cormat)
  }
```

```{r}
# More heatmap
heat_df <- df %>% 
    select(-c(soundname, Filename, Study.y, Language, Diagnosis.y, X)) %>%
    na.omit() %>% 
    mutate(Trial = as.numeric(as.factor(Trial))) %>% 
    mutate(Diagnosis.x = as.numeric(as.factor(Diagnosis.x))) %>% 
    mutate(Gender = as.numeric(as.factor(Gender)))

colnames(heat_df)[7] <- 'AtrPhon'
colnames(heat_df)[6] <- 'SpeechRate'
colnames(heat_df)[8] <- 'ASD_Speak'
```

```{r}
#Building heatmap of correlations
heatmap <- round(cor(heat_df), 2) %>%
    reorder_cormat() %>%
    get_upper_tri() %>%
    melt(na.rm = T) %>%
    ggplot(aes(Var2, Var1, fill = value)) +
    geom_tile(color = "white") +
    scale_fill_gradient2(
        low = "blue",
        high = "red",
        mid = "white",
        midpoint = 0,
        limit = c(-1, 1),
        space = "Lab",
        name = "Pearson\nCorrelation"
    ) +
    theme_minimal() +
    theme(axis.text.x = element_text(
        angle = 45,
        vjust = 1,
        size = 10,
        hjust = 1
    )) +
    coord_fixed() + geom_text(aes(Var2, Var1, label = value),
                              color = "black",
                              size = 3) +
    theme(
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        panel.grid.major = element_blank(),
        panel.border = element_blank(),
        panel.background = element_blank(),
        axis.ticks = element_blank(),
        legend.justification = c(1, 0),
        legend.position = c(0.6, 0.7),
        legend.direction = "horizontal"
    ) +
    guides(fill = guide_colorbar(
        barwidth = 7,
        barheight = 1,
        title.position = "top",
        title.hjust = 0.5
    ))

heatmap
```

## Now we can analyze the data
If you were to examine the meta analysis you would find that the differences (measured as Hedges' g, very close to Cohen's d, that is, in standard deviations) to be the following
- pitch variability (lower, Hedges' g: -0.55, 95% CIs: -1.06, 0.09)
- proportion of spoken time (lower, Hedges' g: -1.26, 95% CIs: -2.26, 0.25)
- speech rate (slower, Hedges' g: -0.75, 95% CIs: -1.51, 0.04)
- pause duration (longer, Hedges' g: 1.89, 95% CIs: 0.72, 3.21). (Duration - Spoken Duration) / PauseN

We need therefore to set up 4 models to see how well our results compare to the meta-analytic findings (Feel free of course to test more features)
Describe the acoustic profile of a schizophrenic voice
*Note* in this section you need to describe the acoustic profile of a schizophrenic voice and compare it with the meta-analytic findings (see 2 and 3 in overview of part 1).

N.B. the meta-analytic findings are on scaled measures. If you want to compare your results with them, you need to scale your measures as well: subtract the mean, and divide by the standard deviation.
N.N.B. We want to think carefully about fixed and random effects in our model. In particular: how should study be included? 
Does it make sense to have all studies put together? Does it make sense to analyze both languages together? Relatedly: does it make sense to scale all data from all studies together?
N.N.N.B. If you want to estimate the studies separately, you can try this syntax:  Feature ~ 0 + Study + Study:Diagnosis + [your randomEffects]. Now you'll have an intercept per each study (the estimates for the controls) and an effect of diagnosis per each study

### NOTES
We only have the Danish studies, if we had had mandarin also we would want to scale them individually. We assume the two studies to be from the same population. Balanced well on diagnosis but not on gender at all. 

*Argument about gender imbalance*: skizo more common in boys, so more males in general. 
```{r}
variable.names(sounddata)

# Make response as a facotr
sounddata$Diagnosis.y <- as.factor(sounddata$Diagnosis.y)

# 1st model: pitch variability: IQR. QUESTION: predict with gender?
    m1 <- glmer(Diagnosis.y ~ scale_IQR + (1|Participant), sounddata, family = "binomial")

# 2nd model: proportion of spoken time: Phonationtime (s): duration of the recording where speech is present
    m2 <- glmer(Diagnosis.y ~ scale_phonation_time + (1|Participant), sounddata, family = "binomial")

# 3rd model: speech rate: (nsyll/dur): average number of syllables per second
    m3 <- glmer(Diagnosis.y ~ scale_speechrate_syll_dur + (1|Participant), sounddata, family = "binomial")

# 4th model: pause duration: how long the pauses last 
    m4 <- glmer(Diagnosis.y ~ scale_pause_dur + (1|Participant), sounddata, family = "binomial")

    summary(m4)
```

PLOTS
```{r}
# ggline(sounddata,
#        x = "Diagnosis.y",
#        y = "pause_dur",
#        col = "Diagnosis",
#        add = c("mean_se", "dodge"),
#        palette = "jco")

```

- Bonus points: cross-validate the models and report the betas and standard errors from all rounds to get an idea of how robust the estimates are. 

```{r}
   # set.seed(1)
   # x <- runif(7)
   # 
   # # Manually scaling
   # (x - mean(x)) / sd(x)
   # 
   # scale(x)
```

## N.B. Remember to save the acoustic features of voice in a separate file, so to be able to load them next time


## Reminder of the report to write 

Part 1 - Can we find a difference in acoustic features in schizophrenia?

1) Describe your sample 
number of studies, number of participants, age, gender, clinical and cognitive features of the two groups. Furthemore, critically assess whether the groups (schizophrenia and controls) are balanced. N.B. you need to take studies into account.


2) Describe the acoustic profile of a schizophrenic voice: which features are different?
E.g. People with schizophrenia tend to have high-pitched voice, and present bigger swings in their prosody than controls. N.B. look also at effect sizes. How do these findings relate to the meta-analytic findings?

3) Discuss the analysis necessary to replicate the meta-analytic findings
Look at the results reported in the paper (see meta-analysis in the readings) and see whether they are similar to those you get.
    3.1) Check whether significance and direction of the effects are similar
    3.2) Standardize your outcome, run the model and check whether the beta's is roughly matched (matched with hedge's g)
which fixed and random effects should be included, given your dataset? E.g. what about language and study, age and gender? Discuss also how studies and languages should play a role in your analyses. E.g. should you analyze each study individually? Or each language individually? Or all together? Each of these choices makes some assumptions about how similar you expect the studies/languages to be.


- Your report should look like a methods paragraph followed by a result paragraph in a typical article (think the Communication and Cognition paper)

```{r}

# VICTOR STUFF
# ?glmer()
# 
# l <- glmer(Diagnosis.y ~ scale_pause_dur + (1|Participant), sounddata, family = "binomial", control = glmerControl(optimizer ="nloptwrap", calc.derivs = FALSE))

```


Rerun: 

First look at the missing data: we should exclude all recordings for which we do not have complete data.
Then count the participants and recordings by diagnosis, report their gender, age and symptom severity (SANS, SAPS and Social)
Finally, do the same by diagnosis and study, to assess systematic differences in studies.
I like to use group_by() %>% summarize() for quick summaries

```{r}
#df <- read.csv('pitch_data_withNA.csv')

# variable.names(df)
# 
# df <- df %>% select(
#     c(
#         Diagnosis.x,
#         Diagnosis.y,
#         Participant,
#         SANS,
#         SAPS,
#         Gender,
#         Study.y,
#         Trial,
#         npause,
#         dur..s.,
#         phonationtime..s.,
#         IQR,
#         speechrate..nsyll.dur.,
#         Age,
#         Mean,
#         Median
#     )
# ) %>%
#     na.omit()
```

# Reading in NEW VEGANERPARTIET CAME WITH COOKIES THIS DAY AND MADE US REALIZE THIS

```{r}
# Looking at npause, some of the data must be spuriously collected, because many of the instances of npause say 0 when duration and phonation time are NOT equal. These have to be sorted out:
map1 <- df$dur..s.== df$phonationtime..s.
map2 <- df$npause == 0

df$npause <- ifelse(map1 == map2, df$npause, NA)

df <- df %>% na.omit()
          
# Pause_dur
df <- df %>% 
    mutate(pause_dur = (dur..s.- phonationtime..s.)/npause)

df$pause_dur <- ifelse(df$npause == 0,0, df$pause_dur) %>% as.numeric()

# Proportion of spoken time
df <- df %>% 
    mutate(prop_spoken_time = phonationtime..s./dur..s.)

# Scaling
df <- df %>% 
    mutate(scale_pause_dur = scale(pause_dur),
           scale_phonation_time = scale(phonationtime..s.),
           scale_IQR = scale(IQR),
           scale_speechrate_syll_dur = scale(speechrate..nsyll.dur.),
           scale_prop_spoken_time = scale(prop_spoken_time),
           scale_mean = scale(Mean))

# The final dataframe is called npause_fixed.csv and has 1424 observations. Gonna read that in to be sure:
df <- read.csv("npause_fixed.csv")
```

# Visualizing
```{r}

df %>% ggplot(aes(Diagnosis.y, Mean_scaled, color = Gender)) + geom_jitter() + labs(x = "Diagnosis", y = "Mean (scaled)")

df %>% ggplot(aes(Gender, Mean_scaled, color = Diagnosis.y)) + geom_jitter() + labs(x = "Gender", y = "Mean (scaled)")

```

# Removing outliers
Removing all values above 3 z-scores
```{r}
df <- df %>% filter(Mean_scaled < 3) 

# plot to see difference
df %>% ggplot(aes(Diagnosis.y, Mean_scaled, color = Gender)) + geom_jitter() + labs(x = "Diagnosis", y = "Mean (scaled)")

df %>% ggplot(aes(Gender, Mean_scaled, color = Diagnosis.y)) + geom_jitter() + labs(x = "Diagnosis", y = "Mean (scaled)")

```

# MODELS 

```{r}

variable.names(df)
# Make response as a factor
df$Diagnosis.y <- as.factor(df$Diagnosis.y)

# 1st model: pitch variability: IQR. QUESTION: predict with gender?
    m1 <- glmer(Diagnosis.y ~ IQR_scaled + (1|Participant), df, family = "binomial", control = glmerControl(optimizer ="nloptwrap", calc.derivs = FALSE))

    summary(m1)
# 2nd model: proportion of spoken time: Phonationtime (s): duration of the recording where speech is present
    m2 <- glmer(Diagnosis.y ~ ProportionSpokenTime_scaled + (1|Participant), df, family = "binomial", control = glmerControl(optimizer ="nloptwrap", calc.derivs = FALSE))

    summary(m2)
# 3rd model: speech rate: (nsyll/dur): average number of syllables per second
    m3 <- glmer(Diagnosis.y ~ SpeechRate_scaled + (1|Participant), df, family = "binomial", control = glmerControl(optimizer ="nloptwrap", calc.derivs = FALSE))

    summary(m3)
# 4th model: pause duration: how long the pauses last 
    m4 <- glmer(Diagnosis.y ~ pause_dur_scaled + (1|Participant), df, family = "binomial", control = glmerControl(optimizer ="nloptwrap", calc.derivs = FALSE))
    
    summary(m4)
    
# Predict diagnosis by mean pitch
    mean_m <- glmer(Diagnosis.y ~ Mean_scaled + (1|Participant), df, family = "binomial", control = glmerControl(optimizer ="nloptwrap", calc.derivs = FALSE))

    summary(mean_m)

```


### HEDGES G
Effect size tells you how much one group differs from another — usually a difference between an experimental group and control group.

So our experimental group is our Schizophrenic people, and our control group is Control people. 

A g of 1 indicates the two groups differ by 1 standard deviation, a g of 2 indicates they differ by 2 standard deviations, and so on. Standard deviations are equivalent to z-scores (1 standard deviation = 1 z-score).

## Now we can analyze the data
If you were to examine the meta analysis you would find that the differences (measured as Hedges' g, very close to Cohen's d, that is, in standard deviations) to be the following
- pitch variability (lower, Hedges' g: -0.55, 95% CIs: -1.06, 0.09) # IQR
- proportion of spoken time (lower, Hedges' g: -1.26, 95% CIs: -2.26, 0.25) #Prop
- speech rate (slower, Hedges' g: -0.75, 95% CIs: -1.51, 0.04) #Speechrate
- pause duration (longer, Hedges' g: 1.89, 95% CIs: 0.72, 3.21). (Duration - Spoken Duration) / PauseN #Pause_dur

Manual way:
esc_mean_sd(grp1m = 10.3, grp1sd = 2.5, grp1n = 60,
grp2m = 12.3, grp2sd = 3.1, grp2n = 56, es.type = "g")

Cohens d way:
```{r}

# To use cohen's d command, we will make two subset df's separately only having Schizophrenia participants or control
df_Sch <- df %>% 
  filter(Diagnosis.y == "Schizophrenia")

df_Con <- df %>% 
  filter(Diagnosis.y == "Control")

# IQR
cohen.d(df_Sch$IQR_scaled, df_Con$IQR_scaled, na.rm = TRUE, pooled = TRUE, paired = FALSE, hedges = TRUE) 
# Our g: -0.26, 95% CIs: -0.36, -0.16
# Meta g: -0.55, 95% CIs: -1.06, 0.09

# Prop of spoken time
cohen.d(df_Sch$ProportionSpokenTime_scaled, df_Con$ProportionSpokenTime_scaled, na.rm = TRUE, pooled = TRUE, paired = FALSE, hedges = TRUE)
# Our g: -0.10, 95% CIs: -0.20, -0.03
# Meta g: -1.26, 95% CIs: -2.26, 0.25

# Speechrate 
cohen.d(df_Sch$SpeechRate_scaled, df_Con$SpeechRate_scaled, na.rm = TRUE, pooled = TRUE, paired = FALSE, hedges = TRUE)
# Our g: -0.25, 95% CIs: -0.35, -0.14
# Meta g: -0.75, 95% CIs: -1.51, 0.04

# Pause duration 
cohen.d(df_Sch$pause_dur_scaled, df_Con$pause_dur_scaled, na.rm = TRUE, pooled = TRUE, paired = FALSE, hedges = TRUE)
# Our g: 0.18, 95% CIs: 0.07, 0.29
# Meta g: 1.89, 95% CIs: 0.72, 3.21

```

Power 
```{r}

```

Confusion matrix
```{r}
# Split the data
NROW(df) 1416


groups <-
  df %>%
  select(Participant) %>%
  distinct(Participant) %>%
  rowwise() %>%
  mutate(group = sample(
    c("train", "test"),
    1,
    replace = TRUE,
    prob = c(0.5, 0.5) # Set weights for each group here
  ))

df_g <- merge(df,groups, by = "Participant")

df_g_train <- (filter(df_g, group == "train")) #720 subs
df_g_test <- (filter(df_g, group == "test")) #696 subs

# Here is some code for checking that all variables in both train and test contain the same levels.
# Look at levels of each variable in each dataframe to find which variable has unequal levels
count_levels <- function(x) factor(x) %>% levels %>% n_distinct

df_g_train %>% summarise_all(count_levels) == df_g_test %>% summarise_all(count_levels)

# Collect all levels of the variable from both datasets and combine into one vector
complete_factor_levels <- c(df_g_train$Diagnosis.y %>% factor %>% levels, df_g_test$Diagnosis.y %>% factor %>% levels) %>% unique

# Assign the complete factor levels to the variable in both datasets
df_g_train <- df_g_train %>% 
  mutate(Diagnosis.y = factor(Diagnosis.y, levels = complete_factor_levels))

df_g_test <- df_g_test %>% 
  mutate(Diagnosis.y = factor(Diagnosis.y, levels = complete_factor_levels))

#Confirm that levels are now the same between the datasets
levels(df_g_train$Diagnosis.y) == levels(df_g_test$Diagnosis.y)

# Fit model to train data
trained_model <- glmer(Diagnosis.y ~ pause_dur_scaled + (1|Participant) , df_g_train, family = "binomial", control = glmerControl(optimizer = "nloptwrap", calc.derivs = FALSE))

# Predict values in test data (type = response gives predicted probabilities)
predicted_probs_test = predict(trained_model, df_g_test, allow.new.levels = TRUE, type = "response")

# Extract actual Diagnosis Schizo categories, i.e. 'true answers' from the original dataframe
actual_categories_test = df_g_test$Diagnosis.y

# Make a dataframe to see predicted probabilities of Diagnosis Schizo category against the actual category
pred_df_test <- tibble(predicted_probs_test, actual_categories_test)

# Make a new column to see more easily what our model predicted: if probability is less than 0.5 - it predicted 'Control' (0), otherwise - it predicted 'Schizophrenia' (1)
pred_df_test$predicted_category = ifelse(pred_df_test$predicted_probs_test < 0.5, "0", "1")

# Let's see first 6 rows
head(pred_df_test)

#make sure predicted category is a factor (at the moment it's a character variable due to the way we specified ifelse outcomes)
pred_df_test$predicted_category <- as_factor(pred_df_test$predicted_category)

pred_df_test$actual_categories_test <- as_factor(pred_df_test$actual_categories_test)

#make the confusion matrix
confusionMatrix(pred_df_test$predicted_category, pred_df_test$actual_categories_test, positive = "1")

```


















MODELS STUFF
```{r}

# df <- df %>% 
#     mutate(scale_SAPS = scale(SAPS))
# 
# m5 <- glmer(
#     Diagnosis.y ~ scale_pause_dur + 
#         scale_IQR + 
#         scale_prop_spoken_time + 
#         scale_speechrate_syll_dur + 
#         (1|Participant), df, 
#     family = "binomial", control = glmerControl(optimizer ="nloptwrap", calc.derivs = FALSE))
# 
# summary(m5)

```

# PLOTS THAT DOESNT WORK RIGHT...
```{r}
# 
# ggline(df,
#        x = "Diagnosis.y",
#        y = "IQR",
#        col = "Diagnosis.y",
#        add = c("mean_se", "dodge"),
#        palette = "jco")
# 
```

