---
title: "L6"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
pacman::p_load(lme4,tidyverse, purrr,dplyr, caret,e1071)
```

# WRITING A FUNCTION notes
You give it a name
Use the function and then the argument
name of function = function(argument) {what you want the function to do}

```{r}
getwd()
pitch_d <- list.files(path = "/Users/pernillebrams/Desktop/EM3/methods3_A3/Pitch/", pattern = ".txt") #list files we want to deal with. This list has all the names of the pitch files
```

# Working on the first file:
```{r}
# Use read-delim function to read it.
df <- read_delim(paste0("/Users/pernillebrams/Desktop/EM3/methods3_A3/Pitch/", pitch_d[1]), delim = "\t") #taking the [1] (first) file
  
# FUNCTION INFORMATION: Paste0 pastes things together with 0 space between the two strings

vars <- str_match(pitch_d[1], "Study(\\d+)D([01])S(\\d+)T(\\d+)") #we want to work with the name of the FIRST file, so [1] in the pitch_d list. \\d+ is a digit. This is really cool, we extract the variables. We specify what we want.

# We can make vars into a dataframe and rename the columns
vars <- as.data.frame(t(vars[2:length(vars)])) # why the t? 
names(vars) <- c("study", "diagnosis", "subject", "trial")

# We want now to extract the descriptors
mean <- mean(df$f0)
sd <- sd(df$f0)
min <- min(df$f0)
max <- max(df$f0)
iqr <- IQR(df$f0)

# We want those not in vectors but as a dataframe, so we add them with cbind
df <- cbind(vars, data.frame(mean,sd,min,max,iqr))
```

Writing the function that we will call read_pitch and calling our argument filename. Put all the fun stuff in the curlys
```{r}
read_pitch <- function(filename){
df <- read_delim(paste0("/Users/pernillebrams/Desktop/EM3/methods3_A3/Pitch/", filename), delim = "\t")
vars <- str_match(pitch_d[1], "Study(\\d+)D([01])S(\\d+)T(\\d+)")
vars <- as.data.frame(t(vars[2:length(vars)])) # why the t? 
names(vars) <- c("study", "diagnosis", "subject", "trial")
mean <- mean(df$f0)
sd <- sd(df$f0)
min <- min(df$f0)
max <- max(df$f0)
iqr <- IQR(df$f0)
df <- cbind(vars, data.frame(mean,sd,min,max,iqr))
return (df) 
}
```
# Applying function 'read_pitch' to each file in the list with map_df:
```{r}

pitch_data = list.files(path = "/Users/pernillebrams/Desktop/EM3/methods3_A3/Pitch/", pattern = ".txt") %>% 
  purrr::map_df(read_pitch)

```

# Making logit 2 prob function:
```{r}
# Making Logit 2 prob function
logit2prob <- function(logit){
  odds <- exp(logit)
  prob <- odds/(1+odds)
  return(prob)
}

# GETTING SOME PROBABILITIES: 

# Probability of Schizophrenia when predictor = 0, that is, when pause_dur is equal to zero:
logit2prob(-13.2417) #Prob of schizo: 1.775015e-06

# Probability of Schizophrenia when pause_dur is increased 1 unit:
logit2prob(0.2045) #Prob of schizo: 0.55%
```


# WORKING WITH OUR OWN DATA
```{r}
# Reading in our own data :) We only have the 4 Danish studies fyi
df <- read.csv("npause_fixed.csv")

df$Diagnosis <- as.factor(df$Diagnosis.y)
df$Participant <- as.character(df$Participant)

# Making a model
pausedur_model <- glmer(
  Diagnosis ~ pause_dur_scaled + (1|Participant), family = "binomial", data = df
)

summary(pausedur_model)

# Making predicted values
df$pred <- logit2prob(predict(pausedur_model))
df$pred <- as.numeric(df$pred)

# Predicted values from the model, and applying function to get the probabilities
df$pred[df$pred > 0.5] = "Schizophrenia"
df$pred[as.numeric(df$pred) <= 0.5] = "Control"
df$pred <- as.factor(df$pred)

# Making confusionMatrix
confusionMatrix(data = df$pred, reference = df$Diagnosis, positive = "Schizophrenia")

```

# MAKING A CURVE that is very help such results
```{r}
pacman::p_load(pROC)
rocCurve <- roc(response = df$Diagnosis, predictor = df$pred)
auc(rocCurve) #amazing
ci(rocCurve)
plot(rocCurve, legacy.axes = TRUE) + title("such results very help")
```

